---
title: "Sensitivity Analysis"
subtitle: "E(tive)Lice"
author: "Tim Szewczyk"
format: html
theme: litera
mainfont: Garamond
monofont: Lucida Console
editor: source
code-fold: true
toc: true
number-sections: true
embed-resources: true
bibliography: references.bib
editor_options: 
  chunk_output_type: console
---

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(glue)
library(sevcheck)
theme_set(theme_bw() + theme(panel.grid=element_blank()))
os <- get_os()
dirs <- switch(
  get_os(),
  linux=list(proj=getwd(),
             out=glue("{getwd()}/out/sensitivity")),
  windows=list(proj=rprojroot::find_rstudio_root_file(),
               out=glue("{rprojroot::find_rstudio_root_file()}/out/sensitivity"))
)
dirf(glue("{dirs$proj}/code/fn"), ".R") |> walk(source)
sim.i <- read_csv(glue("{dirs$out}/sim_i.csv"))

if(file.exists(glue("{dirs$out}/sim_dist.rds"))) {
  sim.dist <- readRDS(glue("{dirs$out}/sim_dist.rds"))
} else {
  n_sim <- 1e4
  post_dir <- glue("{dirs$proj}/data/lit/fit")
  egg_post <- list(constant=read_csv(glue("{post_dir}/egg_temp_constant_post.csv"), show_col_type=F),
                   logistic=read_csv(glue("{post_dir}/egg_temp_logistic_post.csv"), show_col_types=F),
                   linear=read_csv(glue("{post_dir}/egg_temp_linear_post.csv"), show_col_types=F)) |>
    map(~.x |> mutate(across(everything(), ~signif(.x, 3))) |>
          unite("all", everything(), sep=",") %>%
          .$all)
  mort_post <- list(constant=read_csv(glue("{post_dir}/mort_sal_constant_post.csv"), show_col_types=F),
                    logistic=read_csv(glue("{post_dir}/mort_sal_logistic_post.csv"), show_col_types=F)) |>
    map(~.x |> mutate(across(everything(), ~signif(.x, 3))) |>
          unite("all", everything(), sep=",") %>%
          .$all)
  sink_post <- read_csv(glue("{post_dir}/sink_sal_linear_post.csv"), show_col_types=F) |>
    mutate(across(everything(), ~.x/1e3)) #mm/s to m/s
  sim.dist <- sample_parameter_distributions(n_sim, dirs$out, "lhs", egg_post, mort_post, sink_post) |>
    mutate(across(any_of(c("swimColdNauplius", "variableDh", "variableDhV")), 
                  ~.x=="true"))
  saveRDS(sim.dist, glue("{dirs$out}/sim_dist.rds"))
}
```

# Parameter ranges

In the sensitivity analysis, all investigated parameters were varied simultaneously across 1,000 random draws from the distributions defined for each. Note that in Kragesteen 2023:

> Here, connectivity is expressed as the total number of particles entering a 3 by 3 grid (each grid is 160 m × 160 m) area around a farm in a day divided by the number of particles released by a farm in a day.

A 3 by 3 grid gives a `{r} 160*3` by `{r} 160*3` m area. If the farm is directly in the center, this corresponds with a NS/EW distance of `{r} 160*3/2` m, or a diagonal distance of `{r} round(sqrt(2*(160*3/2)^2))` m, capturing an area of `{r} round(160^2 * 9)` m^2, assuming no land overlap. A radius of `{r} round(sqrt(160^2*9/pi))` m would be needed for a circle of the same area.

The IP distance assessed here is the radius of a circle centered on the farm location. 

```{r}
#| echo: false

var_pretty <- list("D_h"="Diffusion: horizontal", 
                   "D_hVert"="Diffusion: vertical",
                   "variableDh"="WeStCOMS horizontal diffusion",
                   "variableDhV"="WeStCOMS vertical diffusion",
                   "eggTemp_fn"="Gravid egg production", 
                   "mortSal_fn"="Larval mortality rate", 
                   "swimUpSpeedNaupliusMean"="Upward swimming speed: N", 
                   "swimDownSpeedNaupliusMean"="Downward sinking speed: N",
                   "swimUpSpeedCopepodidMean"="Upward swimming speed: C", 
                   "swimDownSpeedCopepodidMean"="Downward sinking speed: C",
                   "passiveSinkRateSal"="Salinity-dependent sink rate",
                   "salinityThreshNaupliusMin"="Salinity: Lower threshold: N", 
                   "salinityThreshNaupliusMax"="Salinity: Upper threshold: N", 
                   "salinityThreshCopepodidMin"="Salinity: Lower threshold: C", 
                   "salinityThreshCopepodidMax"="Salinity: Upper threshold: C", 
                   "salRangeN"="Salinity range: N",
                   "salRangeC"="Salinity range: C",
                   "lightThreshNauplius"="Light threshold: N", 
                   "lightThreshCopepodid"="Light threshold: C",
                   "swimColdNauplius"="Cold preference: N",
                   "viableDegreeDays"="Development: N",
                   "maxDepth"="Maximum preferred depth",
                   "connectivityThresh"="IP radius") |>
  as_tibble() |>
  pivot_longer(everything(), names_to="var", values_to="var_pretty") |>
  mutate(var_order=factor(var, 
                          levels=c("D_h", "D_hVert", "variableDh", "variableDhV",
                                   "maxDepth", "connectivityThresh", "passiveSinkRateSal",
                                   "eggTemp_fn", "mortSal_fn", "viableDegreeDays",
                                   "swimColdNauplius",
                                   "salinityThreshNaupliusMin", "salinityThreshNaupliusMax",
                                   "salinityThreshCopepodidMin", "salinityThreshCopepodidMax",
                                   "salRangeN", "salRangeC",
                                   "swimDownSpeedNaupliusMean", "swimUpSpeedNaupliusMean",
                                   "swimDownSpeedCopepodidMean", "swimUpSpeedCopepodidMean",
                                   "lightThreshNauplius", "lightThreshCopepodid"))) |>
  arrange(var_order) |>
  mutate(
    units=list(expression(m^2 %.% s^-1), expression(m^2 %.% s^-1), "", "", "m", "m", "",
               "Function", "Function", expression(degree*C %.% d), "",
               "psu", "psu", "psu", "psu",  "psu", "psu", 
               expression(cm %.% s^-1), expression(cm %.% s^-1),
               expression(cm %.% s^-1), expression(cm %.% s^-1),
               expression(mu*'mol' %.%~'m'^-2 %.% s^-1), expression(mu*'mol' %.%~'m'^-2 %.% s^-1)),
    varNumeric=c(1, 1, 0, 0,
                 1, 1, 0, 
                 0, 0, 1, 0, 
                 1, 1, 
                 1, 1,
                 1, 1, 
                 1, 1,
                 1, 1,
                 1, 1),
    var_cor=c(0, 0, 0, 0,
              0, 0, 0, 
              0, 0, 0, 0,
              1, 1, 
              1, 1, 
              1, 1,
              1, 1,
              1, 1,
              1, 1),
    var_pairs=c(1, 2, 3, 4, 
                5, 6, 7,
                8, 9, 10, 11,
                12, 13,
                12, 13,
                14, 14,
                15, 15, 
                15, 15,
                16, 16),
    var_scale=c("log", "log", "binary", "binary",
                "sqrt", "sqrt", "binary", 
                "binary", "binary", "natural", "binary",
                "natural", "natural", 
                "natural", "natural", 
                "natural", "natural", 
                "sqrt", "sqrt",
                "sqrt", "sqrt",
                "sqrt", "sqrt"),
    var_range=list(log10(c(1e-4, 1e1)), log10(c(1e-6, 1e0)), c(FALSE, TRUE), c(FALSE, TRUE),
                   c(10, 300)^0.5, c(15, 300)^0.5, c(FALSE, TRUE),
                   c("constant", "logistic"), c("constant", "logistic"), c(30, 55), c(FALSE, TRUE),
                   c(5, 32), c(20, 32), 
                   c(5, 32), c(20, 32),
                   c(0, 15), c(0, 15),
                   c(5e-5, 1e-2)^0.5*100, c(5e-5, 5e-3)^0.5*100,
                   c(1e-4, 5e-3)^0.5*100, c(1e-4, 1e-2)^0.5*100,
                   c(0.05, 0.5)^0.5, c(2e-6, 2e-4)^0.5)
  )
var_pretty <- var_pretty |>
  filter(! var %in% c("salinityThreshNaupliusMax", "salinityThreshNaupliusMin",
                      "salinityThreshCopepodidMax", "salinityThreshCopepodidMin",
                      "salRangeN", "salRangeC",
                      "passiveSinkRateSal",
                      "eggTemp_fn", "mortSal_fn"))
```

```{r}
#| fig-width: 11
#| fig-height: 15

sim.used <- sim.i |>
  mutate(across(starts_with("swimDown"), ~.x*100)) |>
  mutate(across(starts_with("swimUp"), ~-.x*100)) |>
  mutate(salRangeN=salinityThreshNaupliusMax - salinityThreshNaupliusMin,
         salRangeC=salinityThreshCopepodidMax - salinityThreshCopepodidMin)
sim.distr <- sim.dist |>
  mutate(across(starts_with("swimDown"), ~.x*100)) |>
  mutate(across(starts_with("swimUp"), ~-.x*100)) |>
  mutate(salRangeN=salinityThreshNaupliusMax - salinityThreshNaupliusMin,
         salRangeC=salinityThreshCopepodidMax - salinityThreshCopepodidMin)
p_ls <- vector("list", nrow(var_pretty))
for(i in seq_along(p_ls)) {
  if(var_pretty$varNumeric[i]) {
    if(var_pretty$var_scale[i] == "natural") {
      x_scale <- scale_x_continuous()
    } else if (var_pretty$var_scale[i] == "sqrt") {
      x_scale <- scale_x_continuous()
    } else if (var_pretty$var_scale[i] == "log") {
      x_scale <- scale_x_log10(guide="axis_logticks", labels=scales::label_log())
    }
    p_ls[[i]] <- ggplot(sim.distr, aes(.data[[var_pretty$var[i]]])) +
      # geom_density(trim=T, bounds=var_pretty$var_range[[i]]) +
      # geom_histogram(aes(y=after_stat(count)/sum(after_stat(count))),
                     # colour="grey30", bins=30, closed="right") +
      ggdist::geom_dots(layout="weave", side="top", colour="grey30", fill="grey30") +
      geom_rug(data=sim.used, aes(.data[[var_pretty$var[i]]]), sides="b", alpha=0.05) +
      x_scale +
      labs(subtitle=var_pretty$var_pretty[i],
           x=var_pretty$units[[i]],
           y="Probability") +
      theme(axis.title.y=element_blank())
  } else {
    p_ls[[i]] <- ggplot(sim.distr, aes(.data[[var_pretty$var[i]]])) +
      geom_bar(aes(y=after_stat(count)/sum(after_stat(count))), 
               colour="grey30") +
      geom_point(data=sim.used |> group_by(.data[[var_pretty$var[i]]]) |> 
                   summarise(N=n()) |> ungroup() |> mutate(p=N/sum(N)),
                 aes(y=p), shape=3) +
      labs(subtitle=var_pretty$var_pretty[i],
           x=var_pretty$units[[i]],
           y="Probability") + 
      ylim(0, 1) +
      theme(axis.title.y=element_blank())
  }
}
sim.distrSub <- sim.distr |>
  sample_n(5000)
p_ls[[length(p_ls)+1]] <- sim.distrSub |>
  select(i, starts_with("salinity")) |>
  mutate(salinity=list(seq(5, 33, by=0.1))) |>
  unnest(salinity) |>
  mutate(pSink=pmin(pmax((salinityThreshNaupliusMax - salinity) / 
                           (salinityThreshNaupliusMax - salinityThreshNaupliusMin), 
                         0), 1)) |>
  ggplot() + 
  geom_line(aes(salinity, pSink, group=i), alpha=0.02) +
  geom_rug(data=sim.distr, aes(salinityThreshNaupliusMax), alpha=0.05, sides="b") +
  geom_rug(data=sim.distr, aes(salinityThreshNaupliusMin), alpha=0.05, sides="t") +
  labs(subtitle="Sinking probability: N",
       x="Salinity (psu)",
       y="Probability") +
  theme(legend.position="none")
p_ls[[length(p_ls)+1]] <- sim.distrSub |>
  select(i, starts_with("salinity")) |>
  mutate(salinity=list(seq(5, 33, by=0.1))) |>
  unnest(salinity) |>
  mutate(pSink=pmin(pmax((salinityThreshCopepodidMax - salinity) / 
                           (salinityThreshCopepodidMax - salinityThreshCopepodidMin), 
                         0), 1)) |>
  ggplot() + 
  geom_line(aes(salinity, pSink, group=i), alpha=0.02) +
  geom_rug(data=sim.distr, aes(salinityThreshCopepodidMax), alpha=0.05, sides="b") +
  geom_rug(data=sim.distr, aes(salinityThreshCopepodidMin), alpha=0.05, sides="t") +
  labs(subtitle="Sinking probability: C",
       x="Salinity (psu)",
       y="Probability") +
  theme(legend.position="none")
p_ls[[length(p_ls)+1]] <- sim.distrSub |>
  select(i, starts_with("passive")) |>
  mutate(salinity=list(seq(0, 32, by=0.1))) |>
  unnest(salinity) |>
  mutate(sinkRate=100*(passiveSinkInt + passiveSinkSlope*salinity)) |>
  ggplot(aes(salinity, sinkRate, group=i, colour=passiveSinkRateSal)) + 
  geom_line(alpha=0.02) +
  scale_colour_manual("Salinity-dependent\nsink rate", values=c("#762a83", "#1b7837")) +
  guides(colour=guide_legend(override.aes=list(alpha=0.8, linewidth=1))) +
  labs(subtitle="Sink rate",
       x="Salinity (psu)",
       y=expression(cm %.% s^-1)) +
  theme(legend.position="none")
p_ls[[length(p_ls)+1]] <- sim.distrSub |>
  select(i, starts_with("eggTemp")) |>
  separate_wider_delim(eggTemp_b, ",", names=paste0("b", 1:4), too_few="align_start") |>
  mutate(across(starts_with("b"), as.numeric)) |>
  mutate(Temp=list(seq(5, 20, by=0.1))) |>
  unnest(Temp) |>
  mutate(eggs=if_else(eggTemp_fn=="constant",
                      b1,
                      b1 / (1+exp(-b2*(Temp-b3))) + b4)) |>
  ggplot() + 
  geom_line(aes(Temp, eggs, group=i, colour=eggTemp_fn), alpha=0.02) +
  scale_colour_manual(values=c("#762a83", "#1b7837")) +
  guides(colour=guide_legend(override.aes=list(alpha=0.8, linewidth=1))) +
  labs(subtitle="Egg production",
       x="Temperature (C)",
       y="Eggs per adult female") +
  theme(legend.position="none")
p_ls[[length(p_ls)+1]] <- sim.distrSub |>
  select(i, starts_with("mortSal")) |>
  separate_wider_delim(mortSal_b, ",", names=paste0("b", 1:4), too_few="align_start") |>
  mutate(across(starts_with("b"), as.numeric)) |>
  mutate(salinity=list(seq(5, 33, by=0.1))) |>
  unnest(salinity) |>
  mutate(pMort=if_else(mortSal_fn=="constant",
                      b1,
                      b1 / (1+exp(-b2*(salinity-b3))) + b4)) |>
  ggplot() + 
  geom_line(aes(salinity, pMort, colour=mortSal_fn, group=i), alpha=0.02) +
  scale_colour_manual(values=c("#762a83", "#1b7837")) +
  guides(colour=guide_legend(override.aes=list(alpha=0.8, linewidth=1))) +
  labs(subtitle="Hourly mortality rate",
       x="Salinity (psu)",
       y="Mortality rate") +
  theme(legend.position="none")

cowplot::plot_grid(plotlist=p_ls, nrow=5, ncol=4, align="hv", axis="trbl")
```



# Smolt migration: Spring IP

## Overall

## By location

### Variability

### Maps: Overall mean, variability

### Characteristics

## Variable Importance

### Overall

### By location 

### Maps

### Accumulated Local Effects


# Farm infestations: Late summer - autumn IP

## Overall 

## By location

### Variability

### Map

### Characteristics

## Variable Importance

## Accumulated Local Effects





# Farm connectivity

How much do connectivity networks change across parameterizations?
