---
title: "E(tive)Lice verification data simulation"
author: "Tim Szewczyk"
format: html
theme: litera
mainfont: Garamond
monofont: Lucida Console
fontsize: 10pt
page-layout: full
editor: visual
toc: true
number-sections: true
fig-responsive: true
embed-resources: true
warning: false
message: false
bibliography: references.bib
editor_options: 
  chunk_output_type: console
---

# Overview

This document generates data for verification of the joint population model (JPM). The JPM is a multilevel Bayesian model that combines copepodid infestation pressure (IP) from a set of particle tracking simulation variants, calculates ensemble IP as a weighted average across the simulation variants, estimates attachment rates, and models the on-fish progression of daily lice cohorts through time. Observed counts of each stage of lice each week are used as the response variable, with all model parameters estimated jointly. The JPM will be verified using data that is simulated using identical processes with known, realistic parameters to assess the ability of the JPM to accurately and precisely recover parameters.

This script can be modified to use final `biotracker` simulations, best available farm data and the final parameter sets. For the moment, it uses daily averages from `etive28` at each farm with a sample of `biotracker` simulations, running from 2023-01-01 to 2024-12-31.

Releases and infestation pressure are currently being performed by pen rather than by farm. The final version may still use farm-level aggregation. However, the IP catchment area is more in line with reality by using a smaller radius (\~30m) around each pen as identified via satellite imagery compared to a larger radius (\~100m) around a reported point location. Since layouts tend to be 2 x n, a circular buffer fails to capture the correct area. For IP, `biotracker` internally sums the density of lice for all particles entering within the radius of each potential destination for each 120 s substep.

Parameters are based on previous literature, e.g., [@kragesteen2023; @kragesteen2021; @kragesteen2019; @kragesteen2018; @samsing2015; @samsing2016], or quick trial-and-error tuning to get plausible values.

The final output is stored in a directory created within `./data/sim` along with all necessary parameter settings and metadata.

# Setup

```{r}
#| results: false
#| label: "setup1"

library(tidyverse)
library(sf)
library(glue)
library(sevcheck)
proj_dir <- rprojroot::find_rstudio_root_file()
theme_set(theme_bw())
dir(glue("{proj_dir}/code/fn"), ".R", full.names=T) |> walk(source)
save_figs <- F

# farm_order <- c("FFMC27", "SAR1", "APT1", "FFMC32", "FFMC84") # etive only
farm_order <-  c("GORS1", "ARDG1", "FFMC84", "FFMC32", "CALL1", "APT1", "SAR1",
                 "KING1", "SHUI1", "FFMC60", "FFMC41", "FFMC27", "FFMC56")

# mesh footprints
mesh_dir <- ifelse(get_os()=="windows",
                   glue("{proj_dir}/../03_packages/WeStCOMS/data"),
                   "~/hydro/meshes")
wce_fp <- st_read(glue("{mesh_dir}/WeStCOMS2_etive28_meshFootprint.gpkg"))
etive_fp <- st_read(glue("{mesh_dir}/etive28_mesh_footprint.gpkg"))
```

```{r}
#| label: "setup2"
# generate simulation number
sim_num <- str_pad(length(dir(glue("{proj_dir}/data/sim"), "sim"))+1, 2, 'left', '0')
out_dir <- glue("{proj_dir}/data/sim/sim_{sim_num}")
inputs_dir <- glue("{proj_dir}/data/sim/inputs")
```

```{r}
#| label: "setup3"
seed <- sample(1e6, 1)
set.seed(seed)
```

```{r}
#| label: "setup4"
#| echo: false

cat("Sim", sim_num, "running with seed:", seed)
dir.create(out_dir, recursive=T)
```

## Farm data

`farm_i`: Farm metadata. Location and volume = $\sum{Area_i \times \min(depth_i, 20)}$ for elements $i$ clipped to the 30m radius around each farm. The 20m depth limit and 30m radius corresponds with the connectivity calculations as parameterized in biotracker (*2_run_biotracker.R*).

```{r}
#| label: "pen_df"
pen_df <- read_csv(glue("{proj_dir}/data/pen_sites_linnhe_2023-2024.csv"), 
                   show_col_types=F) |>
  mutate(sepaSite=str_split_fixed(pen, "_", 2)[,1]) |>
  filter(sepaSite %in% farm_order) |>
  group_by(sepaSite) |>
  mutate(nPens=n()) |>
  ungroup()
glimpse(pen_df)
```

```{r}
#| label: "farm_i"
farm_i <- read_csv(glue("{inputs_dir}/farm_sites_2023-2024.csv"), show_col_types=F) |>
  inner_join(read_csv(glue("{inputs_dir}/farm_influx_vols.csv"), show_col_types=F) |>
               filter(sim=="01") |>
               mutate(sepaSite=str_split_fixed(pen, "_", 2)[,1]) |>
               group_by(sepaSite) |>
               summarise(vol=sum(vol)) |>
               ungroup(), 
             by="sepaSite") |>
  mutate(firstDay=sample(1:5, n(), replace=T)) |>
  left_join(pen_df |> group_by(sepaSite) |> summarise(nPens=first(nPens)), 
            by="sepaSite") |>
  mutate(sepaSite=factor(sepaSite, levels=farm_order)) |>
  arrange(sepaSite)
# farm_i <- read_csv(glue("{inputs_dir}/farm_i.csv"), show_col_types=F) |>
#   filter(sepaSite %in% farm_order) |>
#   mutate(sepaSite=factor(sepaSite, levels=farm_order)) |>
#   arrange(sepaSite) |>
#   mutate(firstDay=sample(1:5, n(), replace=T)) |>
#   left_join(pen_df |> group_by(sepaSite) |> summarise(nPens=first(nPens)))
farm_i
```

```{r}
#| label: "farm_map"
#| fig-width: 3
#| fig-height: 6
#| fig-cap: "Sites in and near Loch Etive."
#| fig-dpi: 400
#| echo: false

farm_sf <- st_as_sf(farm_i, coords=c("easting", "northing"), crs=27700) |> 
  st_buffer(dist=100)
ggplot() +
  geom_sf(data=wce_fp, fill="grey90", colour="grey50") +
  geom_sf(data=etive_fp, fill="dodgerblue", colour="grey50", alpha=0.25) + 
  geom_sf(data=farm_sf, colour=NA, fill="red") +
  geom_sf_text(data=farm_sf, aes(label=sepaSite), hjust=0, nudge_x=100, size=2.5) +
  # scale_x_continuous(limits=c(184000, 211000), oob=scales::oob_keep) + 
  # scale_y_continuous(limits=c(732000, 745000), oob=scales::oob_keep) +
  scale_x_continuous(limits=c(180000, 217000), oob=scales::oob_keep) + 
  scale_y_continuous(limits=c(730000, 778000), oob=scales::oob_keep) +
  theme(axis.title=element_blank(),
        axis.text=element_blank())
```

`farm_dat`: Daily interpolation of mean adult female lice per fish, reported fish biomass, estimated number of fish assuming 240 fish per tonne, estimated total adult female lice assuming 240 fish per tonne, area within 100m radius, receiving area ratio, logit-transformed receiving area ratio, and whether lice were sampled on a given day. Receiving area ratio is calculated following @kragesteen2023 as $R(W_i) = \sqrt{\frac{W_i}{DV_r}}$ (recommended Method 3), with biomass $W$ = `actualBiomassOnSiteTonnes * 1000` ($kg$), stocking density $D = 20 kg \cdot m^{-3}$, and receiving area $V_r$ = `vol` ($m^3$). I assume here a depth of 20m for the receiving area. Thus, $R(W_i)$ = `RW` is the proportion of copepodids that successfully attach.

```{r}
#| label: "farm_dat"
mort_weekly <- 0.000878 # mean rate from Szewczyk et al 2024
prod_cycles <- read_csv(glue("{inputs_dir}/farm_dat.csv"), show_col_types=F) |>
  select(sepaSite, date, actualBiomassOnSiteTonnes) |>
  mutate(fishIn=actualBiomassOnSiteTonnes > 0) |>
  group_by(sepaSite) |>
  mutate(fishInID=consecutive_id(fishIn)) |>
  filter(fishIn) |>
  group_by(sepaSite, fishInID) |>
  slice_head(n=1) |> # for consecutive 0s, select last day
  group_by(sepaSite) |>
  mutate(productionCycleNumber=row_number()) |>
  ungroup() |>
  select(sepaSite, date, productionCycleNumber)
farm_dat <- read_csv(glue("{inputs_dir}/farm_dat.csv"), show_col_types=F) |>
  right_join(farm_i |> select(sepaSite, firstDay), 
             by=join_by(sepaSite)) |>
  right_join(pen_df |> select(sepaSite, pen, nPens), 
             by=join_by(sepaSite), 
             relationship="many-to-many") |>
  left_join(prod_cycles, by=c("sepaSite", "date")) |>
  fill(productionCycleNumber) |>
  group_by(pen, productionCycleNumber) |>
  mutate(prodDay=row_number()) |>
  ungroup() |>
  mutate(sepaSite=factor(sepaSite, levels=farm_order),
         fishIn=actualBiomassOnSiteTonnes > 0,
         nFish_est=fishIn * maximumBiomassAllowedTonnes*240/nPens * 
           ((1-mort_weekly)^(1/7))^prodDay,
         total_AF=weeklyAverageAf * nFish_est,
         RW=sqrt(fishIn*maximumBiomassAllowedTonnes/nPens*1000/(20*20*pi*30^2)),
         RW_logit=brms::logit_scaled(RW, lb=-1e-5)) |>
  group_by(sepaSite) |>
  mutate(sampled=date %in% (first(date) + first(firstDay) + 7*(0:500)),
         sampled=sampled & (nFish_est > 0)) |>
  ungroup() |>
  select(-firstDay, -nPens) |>
  mutate(nFishSampled=if_else(sampled, sample(20:25, 1), 0))
glimpse(farm_dat)
```

```{r}
#| results: "hide"
#| echo: false

gc()
```

```{r}
#| label: "farm_dat_fig"
#| fig-width: 15
#| fig-height: 9
#| fig-cap: "Farm data based on reported values. Rug on x-axis shows dates where simulated lice will be sampled."
#| fig-dpi: 400
#| echo: false

farm_dat_gg <- ggplot(farm_dat, aes(x=date)) + 
  geom_rug(aes(colour=sampled), sides="b", linewidth=0.1) +
  scale_colour_manual(values=c(NA, "black"), na.value=NA, guide="none") +
    scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
  ylim(0, NA) + 
  facet_wrap(~sepaSite, nrow=1) +
  theme(axis.title.x=element_blank())
p_ls <- list(farm_dat_gg + geom_line(aes(y=nFish_est, group=pen)),
             farm_dat_gg + geom_line(aes(y=weeklyAverageAf, group=pen)),
             farm_dat_gg + geom_line(aes(y=total_AF, group=pen)),
             farm_dat_gg + geom_line(aes(y=RW, group=pen)))
cowplot::plot_grid(plotlist=p_ls, ncol=1, align="hv")

if(save_figs) {
  cowplot::plot_grid(farm_dat_gg +
                       geom_line(aes(y=nFish_est, group=pen)) +
                       ylab("Number of fish"),
                     farm_dat_gg +
                       geom_line(aes(y=weeklyAverageAf * (nFish_est > 0), group=pen)) +
                       ylab("Mean lice per fish (AF)"),
                     ncol=1, align="hv")
  ggsave("admin/project_meetings/figs_temp/dataExamp_farm_dat.png", width=10, height=4)
}
```

```{r}
#| results: "hide"
#| echo: false

rm(farm_dat_gg)
gc()
```

## Environmental data

From `etive28`, I've calculated daily averages across the top 20m at each farm location to summarise the environmental conditions.

```{r}
#| label: "farm_env"
farm_env <- read_csv(glue("{inputs_dir}/farm_env_hourly.csv"), show_col_types=F) |>
  filter(sepaSite %in% farm_order) |>
  mutate(sepaSite=factor(sepaSite, levels=farm_order),
         date=date(time),
         hour=hour(time)) |>
  mutate(day=as.numeric(as.factor(date))) |>
  inner_join(
    farm_dat |> 
      select(sepaSite, pen, date, nFish_est, RW, RW_logit, sampled, nFishSampled), 
      by=join_by(sepaSite, pen, date)) |>
  arrange(sepaSite, time) |>
  mutate(u=u*100, # cm/s
         v=v*100, # cm/s
         w=w*100, # cm/s
         uv=uv*100, 
         uv_sq=uv^2,
         salinity_m30=salinity - 30) |> # recenter so intercept = high salinity
  mutate(across(c(temperature, u, v, w, uv, salinity), ~c(scale(.x)), .names="{.col}_z"))
glimpse(farm_env)
# for back-transforming z-scores
farm_env_avg <- farm_env |>
  reframe(across(where(is.numeric), ~c(mn=mean(.x), sd=sd(.x)))) |>
  select(-ends_with("_z")) |>
  mutate(metric=c("mean", "sd"))
farm_env_avg
# daily values
farm_env_daily <- farm_env |>
  select(-time, -hour, -elapsedHours) |>
  group_by(sepaSite, pen, day, date) |>
  summarise(across(where(is.numeric), mean),
            sampled=first(sampled)) |>
  ungroup()
```

```{r}
#| label: "farm_env_fig"
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Etive28 or WeStCOMs2 at farm locations. Farms are ordered from lowest salinity (light) to highest (dark)."
#| fig-dpi: 400
#| echo: false

farm_env_daily |>
  select(-ends_with("_z"), -"salinity_m30") |>
  pivot_longer(cols=c("u", "v", "w", "uv", "salinity", "temperature"),
               names_to="variable", values_to="value") |>
  mutate(variable=factor(variable,
                         levels=c("salinity", "temperature", "u", "v", "w", "uv"))) |>
  ggplot(aes(date, value, colour=sepaSite, group=pen)) +
  geom_line() +
  scico::scale_colour_scico_d(palette="devon", end=0.8, direction=1, guide="none") +
  facet_grid(variable~sepaSite, scales="free_y") +
  scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
  labs(y="Daily mean, 0-20m (Etive28, WeStCOMS2)") +
  theme(panel.grid.major.y=element_line(colour="grey85", linewidth=0.2),
        panel.grid.minor.y=element_blank(),
        axis.title.x=element_blank())

if(save_figs) {
  farm_env_daily |>
    select(-ends_with("_z"), -"salinity_m30") |>
    pivot_longer(cols=c("u", "v", "w", "uv", "salinity", "temperature"),
                 names_to="variable", values_to="value") |>
    mutate(variable=factor(variable,
                           levels=c("salinity", "temperature", "u", "v", "w", "uv"))) |>
    filter(variable %in% c("uv", "salinity", "temperature")) |>
    ggplot(aes(date, value, colour=sepaSite, group=pen)) +
    geom_line() +
    scico::scale_colour_scico_d(palette="devon", end=0.8, direction=1, guide="none") +
    facet_grid(variable~sepaSite, scales="free_y") +
    scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
    labs(y="Daily mean, 0-20m (Etive28, WeStCOMS2)") +
    theme(panel.grid.major.y=element_line(colour="grey85", linewidth=0.2),
          panel.grid.minor.y=element_blank(),
          axis.title.x=element_blank())
  ggsave("admin/project_meetings/figs_temp/dataExamp_farm_env.png", width=10, height=6)
  }
```

```{r}
#| results: "hide"
#| echo: false

gc()
```

## Infestation pressure

Copepodid infestation pressure was taken from 6 `biotracker` simulations which randomly sampled from a plausible parameter space. The number of copepodids in the top 20m within a 30m radius of each pen was totaled each day. The number of simulations will increase as testing progresses. The parameters that varied were:

```{r}
#| message: false
#| echo: false
#| eval: false

read_csv(glue("E:/EtiveLice/out/biotracker/sim_i.csv")) |>
  select(-i, -outDir, -ends_with("_b")) |>
  mutate(across(where(is.numeric), ~signif(.x, 3))) |>
  glimpse()
```

```{r}
#| label: "influx_df"
influx_df <- read_csv(glue("{inputs_dir}/influx_hourly.csv"), show_col_types=F) |>
  filter(sepaSite %in% farm_order) |>
  full_join(tibble(time=seq_range(range(farm_env$time), by=3600),
                   sepaSite=farm_order[1], pen=paste0(farm_order[1], "_01"), sim="01"),
            by=join_by(sepaSite, sim, pen, time)) |>
  filter(between(time, min(farm_env$time), max(farm_env$time))) |>
  complete(time, pen, sim, fill=list(influx=0, influx_m3=0)) |>
  mutate(date=date(time),
         day=as.numeric(as.factor(date)),
         hour=hour(time),
         sepaSite=str_split_fixed(pen, "_", 2)[,1],
         sepaSite=factor(sepaSite, levels=farm_order)) |>
  select(time, date, day, hour, sepaSite, pen, sim, influx, influx_m3) |>
  arrange(sim, sepaSite, pen, time)
glimpse(influx_df)
```

```{r}
#| label: "influx_fig1"
#| fig-width: 10
#| fig-height: 10
#| fig-cap: "Daily copepodid infestation pressure by farm."
#| fig-dpi: 400
#| echo: false
#| eval: false
ggplot(influx_df, aes(time, influx_m3, colour=sepaSite, group=pen)) +
  geom_line(linewidth=0.1) +
  scico::scale_colour_scico_d(palette="devon", end=0.8, direction=1, guide="none") +
  scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
  scale_y_continuous("Mean daily copepodids / m3 in top 20m") +
  facet_grid(sim~sepaSite) +
  theme(axis.title.x=element_blank())
```

```{r}
#| label: "influx_fig2"
#| fig-width: 10
#| fig-height: 10
#| fig-cap: "Daily copepodid infestation pressure (4th root) by farm."
#| fig-dpi: 400
#| echo: false
#| eval: false
ggplot(influx_df, aes(time, influx_m3^0.25, colour=sepaSite, group=pen)) +
  geom_line(linewidth=0.1) +
  scico::scale_colour_scico_d(palette="devon", end=0.8, direction=1, guide="none") +
  scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
  scale_y_continuous("(Mean daily copepodids / m3)^0.25") +
  facet_grid(sim~sepaSite) +
  theme(axis.title.x=element_blank())
if(save_figs) {
  ggsave("admin/project_meetings/figs_temp/dataExamp_IP.png", width=10, height=6)
}
```

```{r}
#| label: "influx_fig3"
#| fig-width: 10
#| fig-height: 10
#| fig-cap: "Daily copepodid infestation pressure (log) by farm."
#| fig-dpi: 400
#| eval: false
#| echo: false
ggplot(influx_df, aes(time, log1p(influx_m3), colour=sepaSite, group=pen)) +
  geom_line(linewidth=0.1) +
  scico::scale_colour_scico_d(palette="devon", end=0.8, direction=1, guide="none") +
  scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
  scale_y_continuous("ln(1 + Mean daily copepodids / m3)") +
  facet_grid(sim~sepaSite) +
  theme(axis.title.x=element_blank())
if(save_figs) {
  ggsave("admin/project_meetings/figs_temp/dataExamp_IPlog.png", width=10, height=6)
}
```

The simulations vary quite a lot in the infestation pressure, with considerable variation among pens as well.

## Population settings

::: callout-note
Attachment is modelled using hourly outputs. Current speed and IP both show a lot of intra-day fluctuation and using daily averages may severely misrepresent dynamics depending on whether high IP coincides with high attachment rates or low attachment rates.

Progression on-fish is more reasonable to model with a daily timestep given that 1) the population is largely fixed, 2) mean values will capture overall exposure throughout the day, and 3) salinity and temperature show much less tidal variability (though still some).
:::

```{r}
#| label: "brms"
#| eval: false
#| echo: false
# Generate realistic variability in parameters using Bayesian regression
library(brms)
library(plotly)

# Pr(attach)
# attach_mn <- c(-2.25, .8, 0.2, 0.2, -0.005)
attach_mn <- c(0, 1, 0.2, 0.25, 0.2, -0.1)
attach_sigma <- 0.7
attach_df <- expand_grid(
  salinity_z=seq_range(farm_env$salinity_z, length.out=100),
  uv_z=seq_range(farm_env$uv_z, length.out=100),
  RW=seq_range(filter(farm_env, RW>0)$RW, length.out=100),
  temp_z=seq_range(farm_env$temperature_z, length.out=100)
) |>
  slice_sample(n=50) |>
  mutate(RW_logit=brms::logit_scaled(RW, lb=-1e-5),
         salinity=salinity_z*farm_env_avg$salinity[2] + farm_env_avg$salinity[1],
         temp=temp_z*farm_env_avg$temperature[2] + farm_env_avg$temperature[1]) |>
  mutate(eta=attach_mn[1] + 
           attach_mn[2] * RW_logit + 
           attach_mn[3] * salinity_z + +
           attach_mn[4] * temp_z +
           attach_mn[5] * uv_z +
           attach_mn[6] * uv_z^2,
         obs=eta + rnorm(n(), 0, attach_sigma))
ggplot(attach_df, aes(uv_z, plogis(eta), colour=RW)) + geom_point() + ylim(0, 1)
ggplot(attach_df, aes(temp, plogis(eta), colour=RW)) + geom_point() + ylim(0, 1)
ggplot(attach_df, aes(RW, plogis(eta), colour=uv_z)) + geom_point() + ylim(0, 1) + 
  geom_abline() + scale_colour_viridis_c(option="turbo")
ggplot(attach_df, aes(RW, plogis(eta), colour=temp)) + geom_point() + ylim(0, 1) + 
  geom_abline() + scale_colour_viridis_c(option="turbo")
plot_ly(attach_df, x=~uv_z, y=~RW, z=~salinity, color=~plogis(eta), type="scatter3d")
plot_ly(attach_df, x=~uv_z, y=~RW, z=~temp, color=~plogis(eta), type="scatter3d")
summary(plogis(attach_df$eta))
hist(plogis(attach_df$eta), xlim=c(0,1))


attach_out <- brm(obs ~ 0 + RW_logit + salinity_z + temp_z + uv_z + I(uv_z^2),
                  data=attach_df, cores=4)
as_draws_df(attach_out) |>
  select(starts_with("b_")) |>
  write_csv("data/sim/inputs/posterior_prAttach.csv")

# Pr(surv) 
# surv_mn=cbind(c(-2.5, -1, -1, -1.5),
#               c(0.2, 0.18, 0.18, 0.16))
surv_mn <- cbind(qlogis(c(0.95, 0.98, 0.99, 0.97)),
                 c(0.2, 0.18, 0.18, 0.16))
surv_sigma <- 0.7
surv_df <- expand_grid(
  salinity=seq_range(farm_env$salinity_m30, length.out=100),
  stage=factor(1:4)
) |>
  slice_sample(n=10, by=stage) |> 
  ungroup() |>
  mutate(eta=surv_mn[stage,1] + surv_mn[stage,2]*salinity,
         obs=eta + rnorm(n(), 0, surv_sigma),
         int1=as.numeric(stage==1),
         int2=as.numeric(stage==2),
         int3=as.numeric(stage==3),
         int4=as.numeric(stage==4),
         sal1=as.numeric(stage==1)*salinity,
         sal2=as.numeric(stage==2)*salinity,
         sal3=as.numeric(stage==3)*salinity,
         sal4=as.numeric(stage==4)*salinity)
surv_out <- brm(obs ~ 0 + int1 + int2 + int3 + int4 + sal1 + sal2 + sal3 + sal4,
                 data=surv_df, cores=4)
as_draws_df(surv_out) |>
  select(starts_with("b_")) |>
  write_csv("data/sim/inputs/posterior_prSurv.csv")
```

```{r}
#| label: "set_params"
use_farm_avg <- TRUE # F: use pens
info <- list(nDays=n_distinct(influx_df$date),
             nHours=n_distinct(influx_df$time),
             nFarms=n_distinct(influx_df$pen),
             nSims=n_distinct(influx_df$sim),
             nStages=3,
             nPens=rep(1, n_distinct(influx_df$pen)),
             IP_penVolume=(read_csv(glue("{inputs_dir}/farm_influx_vols.csv"), show_col_types=F) |>
                             filter(sim=="01") |>
                             select(pen, vol) |>
                             mutate(sepaSite=str_split_fixed(pen, "_", 2)[,1]) |>
                             inner_join(farm_i |> select(sepaSite), y=_))$vol,
             dateRange=range(influx_df$date))
params <- list(
  # Ensemble proportional weight for each biotracker simulation
  ensWts_p=rlnorm(info$nSims, sdlog=0.75) |> make_compositional(),
  # Pr(copepodid attachment): RW_logit, salinity, temp, uv, uv^2
  attach_beta=glue("{inputs_dir}/posterior_prAttach.csv") |>
    read_csv() |> sample_n(1) |> unlist(),
  # On-fish survival: chalimus, pre-adult, adult, gravid
  #   logit(p) = f(salinity); intercept, slope by stage
  surv_beta=glue("{inputs_dir}/posterior_prSurv.csv") |>
    read_csv() |> sample_n(1) |> unlist() |> matrix(nrow=2, byrow=T),
  # On-fish survival: sd among farms (Int.) by stage: cleaner fish, other characteristics
  surv_int_farm_sd=abs(brms::rstudent_t(4, 3, 0, 0.75)),
  # Cumulative GDD post-attachment needed to transition to next stage
  thresh_GDD=cbind( 
    runif(3, min=c(125, 325, 525), max=c(175, 375, 575)), # F: Pr/Ad/Gr
    runif(3, min=c(105, 290, 5e3), max=c(155, 340, 5e3)) # M: Pr/Ad/NA
  )[1:(info$nStages-1),],
  lifespan=rnorm(1, 1500, 25),
  # Pr(detect) by stage
  detect_p=c(runif(3, 
                   min=c(0, 0.5, 1),
                   max=c(0.5, 0.99, 1)),
             1)[1:info$nStages],
  # Precision parameter for fish-level lice counts ~negbinom(mu, phi)
  # Distribution tends toward poisson as phi -> Inf
  # nb_prec=abs(rnorm(1, 0, 2)),
  nb_prec=rchisq(1, 2),
  # Background infestation rate from wild fish (copepodids/m3/d)
  IP_bg_m3=(runif(1, min=(1e-5)^0.5, max=(0.1)^0.5))^2,
  # Scaling factor for IP -> attached: from linear to 4th root with mean = sq root 
  #IP_scale=rbeta(n=1, 1.5, 3)*3 + 1, # !!! DEPRECATED !!!
  # Half saturation constant for attachment (cop/m3/day) 
  IP_halfSat_m3=runif(n=1, 1, 50) + rbinom(1, 1, 0.5)*1e6,
  # Treatment threshold for AF
  treat_thresh=sample(c(0.5, 1), 1),
  treat_efficacy=rbeta(1, 4, 2)
)
# Randomly set salinity and uv effects to 0 for attachment
params$attach_beta[2:5] <- params$attach_beta[2:5] * rbinom(3, 1, 0.75)[c(1,2,3,3)]
# Randomly set salinity effects to 0 for survival
params$surv_beta[2,] <- params$surv_beta[2,] * rbinom(4, 1, 0.75)
params$surv_beta <- params$surv_beta[,1:info$nStages]
params$surv_int_farm_sd <- params$surv_int_farm_sd[1:info$nStages]
# Background copepodid count per pen
params$IP_bg <- params$IP_bg_m3*info$IP_penVolume
# Attachment half saturation IP per pen
params$IP_halfSat <- params$IP_halfSat_m3*info$IP_penVolume

mnDaysStage_df <- expand_grid(Ch=seq(125, 175, length.out=100),
            PA=seq(325, 375, length.out=100)) |>
  mutate(PA=PA-Ch) |>
  pivot_longer(everything(), names_to="stage", values_to="GDD") |>
  mutate(temperature=list(seq_range(farm_env$temperature, length.out=100))) |>
  unnest(temperature) |>
  mutate(days=GDD/temperature,
         temp_z=(temperature-farm_env_avg$temperature[1])/farm_env_avg$temperature[2])
params$mnDaysStageCh <- summary(
  lm(days ~ temp_z, data=mnDaysStage_df |> filter(stage=="Ch")))$coefficients[,1:2] * 
  rep(c(1, sqrt(nrow(mnDaysStage_df)/2)), each=2)
params$mnDaysStagePA <- summary(
  lm(days ~ temp_z, data=mnDaysStage_df |> filter(stage=="PA")))$coefficients[,1:2] * 
  rep(c(1, sqrt(nrow(mnDaysStage_df)/2)), each=2)
```

In this particular simulation, the settings are:

```{r}
#| echo: false
map(params, ~signif(.x, 3))
```

```{r}
#| label: "param_fig"
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 8
# attachment
p_prAttach <- expand_grid(
  salinity_z=seq_quantiles(farm_env$salinity_z, 0.05, 0.995, length.out=100),
  uv_z=seq_quantiles(farm_env$uv_z, 0, 0.9995, length.out=100),
  RW=c(0.25, 0.5),
  temp_z=seq_quantiles(farm_env$temperature_z, 0.05, 0.995, length.out=4)
) |>
  mutate(uv_sq=uv_z^2,
         uv=uv_z*farm_env_avg$uv[2] + farm_env_avg$uv[1],
         salinity=salinity_z*farm_env_avg$salinity[2] + farm_env_avg$salinity[1],
         temp=temp_z*farm_env_avg$temperature[2] + farm_env_avg$temperature[1],
         temp_lab=paste(round(temp, 1), " C"),
         temp_lab=factor(temp_lab, levels=unique(temp_lab)),
         RW_logit=brms::logit_scaled(RW, lb=-1e-5),
         RW_lab=paste("RW:", round(RW, 2))) |>
  mutate(pr_attach=plogis(params$attach_beta[1] * RW_logit + 
                            params$attach_beta[2] * salinity_z +
                            params$attach_beta[3] * temp_z +
                            params$attach_beta[4] * uv_z +
                            params$attach_beta[5] * uv_sq)) |>
  ggplot(aes(salinity, uv, fill=pr_attach)) + 
  geom_raster() + 
  # scale_fill_viridis_b("P(attach)", option="turbo", 
  #                      limits=c(0, 1), breaks=seq(0, 1, by=0.05),
  #                      labels=c("0", rep("", 4), "0.25", rep("", 4), "0.5", rep("", 4),
  #                               "0.75", rep("", 4), "1")) +
  scale_fill_viridis_b("P(attach)", option="turbo", 
                       limits=c(0, 1), breaks=seq(0, 1, by=0.1),
                       labels=c("0", rep("", 4), "0.5", rep("", 4), "1")) +
  # scale_fill_viridis_c("Max P(attach)", option="turbo", 
  #                      limits=c(0, 1), breaks=seq(0, 1, by=0.25),
  #                      labels=c("0", "0.25", "0.5", "0.75", "1")) +
  facet_grid(RW_lab~temp_lab) +
  scale_x_continuous("Salinity (psu)", breaks=c(10, 20, 30)) +
  labs(y="Water speed (cm/s)") +
  theme(legend.position="bottom",
        legend.key.width=unit(0.6, "cm"),
        legend.key.height=unit(0.2, "cm"))
p_prAttachIP <- expand_grid(
  IP=seq_quantiles(influx_df$influx_m3, 0, 0.99995, by=0.01)*max(info$IP_penVolume),
  maxPrAttach=seq(0, 1, by=0.01)
  ) |>
  mutate(pr_attach=(maxPrAttach*max(params$IP_halfSat))/(IP + max(params$IP_halfSat))) |>
  ggplot(aes(IP/max(info$IP_penVolume), maxPrAttach, fill=pr_attach)) + 
  geom_raster() +
  annotate("text", x=0.075, y=0.85, parse=T, hjust=0, vjust=0, size=3,
           label=paste0("IP[bg]:~", round(params$IP_bg_m3, 3), "~cop~m^-3~h^-1")) +
  annotate("text", x=0.075, y=0.75, parse=T, hjust=0, vjust=0, size=3,
           label=paste0("P[attach]~halfSat.:~", round(params$IP_halfSat_m3, 1), 
                        "~cop~m^-3~h^-1")) +
  # scale_fill_viridis_b("P(attach)", option="turbo", 
  #                      limits=c(0, 1), breaks=seq(0, 1, by=0.05),
  #                      labels=c("0", rep("", 4), "0.25", rep("", 4), "0.5", rep("", 4),
  #                               "0.75", rep("", 4), "1")) +
  scale_fill_viridis_b("P(attach)", option="turbo", 
                       limits=c(0, 1), breaks=seq(0, 1, by=0.1),
                       labels=c("0", rep("", 4), "0.5", rep("", 4), "1")) +
  # scale_fill_viridis_c("P(attach)", option="turbo", 
  #                      limits=c(0, 1), breaks=seq(0, 1, by=0.25),
  #                      labels=c("0", "0.25", "0.5", "0.75", "1")) +
  ylim(0, 1) +
  labs(x="IP (copepodids/m3/h)", y="Max P(attach)") +
  theme(legend.position="bottom",
        legend.key.width=unit(0.6, "cm"),
        legend.key.height=unit(0.2, "cm"))
# ensemble weights
p_ensWts <- tibble(sim=unique(influx_df$sim), 
                   p=params$ensWts_p) |>
  ggplot() +
  geom_bar(aes(x=sim, y=p, fill=sim), stat="identity", colour="grey30") +
  scale_fill_viridis_d("Sim", option="turbo", guide="none") +
  scale_y_continuous("Ensemble weight p", limits=c(0, 1),
                     breaks=round(seq(0, 1, by=1/info$nSims), 2)) +
  xlab("Biotracker simulation")
# survival
p_prSurv <- expand_grid(
  salinity=seq_range(farm_env$salinity_m30, length.out=100),
  stage=1:info$nStages
) |>
  mutate(pr_surv=plogis(params$surv_beta[1,stage] +
                          params$surv_beta[2,stage]*salinity),
         pr_surv_mSD=plogis(params$surv_beta[1,stage] - params$surv_int_farm_sd[stage] +
                              params$surv_beta[2,stage]*salinity),
         pr_surv_pSD=plogis(params$surv_beta[1,stage] + params$surv_int_farm_sd[stage] +
                              params$surv_beta[2,stage]*salinity),
         stage=factor(stage, 
                      labels=c("Ch", "PA", "Ad", "Gr")[1:info$nStages])) |>
  ggplot(aes(salinity+30, pr_surv, colour=stage, fill=stage)) +
  geom_ribbon(aes(ymin=pr_surv_mSD, ymax=pr_surv_pSD), alpha=0.25, colour=NA) +
  geom_line() +
  scale_colour_manual(values=RColorBrewer::brewer.pal(4, "Paired")[c(1,2,4)]) +
  scale_fill_manual(values=RColorBrewer::brewer.pal(4, "Paired")[c(1,2,4)]) +
  labs(x="Salinity (psu)", y="P(survival) ± farm sd") +
  ylim(0, 1) +
  theme(legend.title=element_blank(),
        legend.position="inside",
        legend.position.inside=c(0.8, 0.15),
        legend.key.height=unit(0.2, "cm"),
        legend.background=element_blank())
# development
p_GDD <- tibble(Sex=rep(c("F", "M"), each=info$nStages),
                Stage=rep(1:info$nStages, times=2),
                GDD=c(rbind(params$thresh_GDD, rep(5e3, 2)))) |>
  mutate(Stage=factor(Stage, 
                      labels=c("Ch", "PA", "Ad", "Gr")[1:info$nStages])) |>
  full_join(expand_grid(Sex=c("M", "F"), 
                        GDD=1:params$lifespan)) |>
  arrange(Sex, GDD) |>
  group_by(Sex) |>
  fill(Stage, .direction="up") |>
  filter(!is.na(Stage)) |>
  ggplot(aes(GDD, Stage, colour=Sex, group=Sex)) +
  geom_line() +
  labs(x="Growing degree days") +
  xlim(0, 1600) +
  theme(axis.title.y=element_blank(),
        legend.title=element_blank(),
        legend.position="inside",
        legend.position.inside=c(0.8, 0.15),
        legend.key.height=unit(0.2, "cm"),
        legend.background=element_blank())
p_daysStage <- tibble(Sex=rep(c("F", "M"), each=info$nStages-1),
                Stage=rep(1:(info$nStages-1), times=2),
                GDD=c(rbind(params$thresh_GDD))) |>
  group_by(Sex) |>
  mutate(GDD=GDD-cumsum(dplyr::lag(GDD, default=0))) |>
  ungroup() |>
  mutate(Stage=factor(Stage, labels=c("Ch", "PA"))) |>
  mutate(temperature=list(seq_range(farm_env$temperature, length.out=100))) |>
  unnest(temperature) |>
  mutate(days=GDD/temperature) |>
  ggplot(aes(temperature, days, colour=Stage, linetype=Sex)) + 
  geom_line() +
  scale_colour_manual("Transition", values=RColorBrewer::brewer.pal(3, "Paired")[1:2]) +
  labs(x="Temperature (C)", y="Stage length (days)") +
  ylim(0, 30) +
  theme(legend.position="inside",
        legend.position.inside=c(0.8, 0.75),
        legend.spacing.y=unit(0.1, "mm"),
        legend.title=element_blank(),
        legend.key.height=unit(0.2, "cm"),
        legend.background=element_blank())
# detection
p_detect <- tibble(Stage=factor(1:info$nStages, 
                                labels=c("Ch", "PA", "Ad", "Gr")[1:info$nStages]),
       pr_detect=params$detect_p) |>
  ggplot(aes(pr_detect, Stage)) + 
  geom_bar(stat="identity", fill="grey", colour="grey30") +
  scale_x_continuous("P(Detect)", limits=c(0, 1)) +
  theme(axis.title.y=element_blank())
# sampling
p_nb <- expand_grid(mu=c(0.01, 0.1, 0.5, 1, 5),
            y=0:10) |>
  mutate(p=dnbinom(y, mu=mu, size=params$nb_prec),
         mu_lab=paste("mu :", mu)) |>
  ggplot(aes(y, p)) + 
  geom_bar(stat="identity", colour="grey30", fill="cadetblue") +
  geom_text(data=tibble(y=8, p=0.9, mu_lab="mu : 0.01", 
                        label=paste0("phi: ", round(params$nb_prec, 2))),
            aes(label=label), parse=T) +
  scale_x_continuous("Observed lice count", breaks=seq(0, 10, by=2)) +
  scale_y_continuous("Probability") +
  facet_grid(.~mu_lab, labeller=label_parsed)
cowplot::plot_grid(cowplot::plot_grid(p_prAttach, p_prAttachIP,
                                      nrow=1, rel_widths=c(1, 0.5)),
                   cowplot::plot_grid(p_ensWts, p_prSurv, p_daysStage, p_detect,
                                      nrow=1, axis="tb", align="h"),
                   p_nb, ncol=1, rel_heights=c(0.9, 0.5, 0.55))
if(save_figs) {
  ggsave("admin/project_meetings/figs_temp/dataExamp_params.png", width=10, height=9)
}
```

```{r}
#| results: "hide"
#| echo: false

gc()
```

# Simulation

The simulation function is wrapped into `simulate_farm_pops()` in `code/fn/simulation.R`. Note that these loops can be vectorized for performance if needed. For forward simulation everything runs very fast, but the Stan model will likely need to be vectorized.

```{r}
#| echo: false
#| eval: false
#| code-fold: true
#| file: fn/simulation.R
```

```{r}
#| label: "use_farm_avg"
#| message: false
if(use_farm_avg) {
  farm_env <- farm_env |>
    group_by(time, date, day, hour, sepaSite) |>
    summarise(pen=first(pen),
              sampled=first(sampled),
              across(all_of(c("RW_logit", "temperature", "temperature_z", 
                              "salinity", "salinity_z", "salinity_m30",
                              "uv", "uv_sq", "uv_z")), mean),
              across(all_of(c("nFish_est", "nFishSampled")), sum)) |>
    ungroup() |>
    mutate(uv_z_sq=uv_z^2)
  farm_env_daily <- farm_env_daily |>
    group_by(date, day, sepaSite) |>
    summarise(pen=first(pen),
              sampled=first(sampled),
              across(all_of(c("RW_logit", "temperature", "temperature_z", 
                              "salinity", "salinity_z", "salinity_m30",
                              "uv", "uv_sq", "uv_z")), mean),
              across(all_of(c("nFish_est", "nFishSampled")), sum)) |>
    ungroup() |>
    mutate(uv_z_sq=uv_z^2)
  influx_df <- influx_df |> 
    group_by(time, date, day, hour, sepaSite, sim) |>
    summarise(pen=first(pen),
              influx=sum(influx)) |>
    ungroup()
  info$nFarms <- n_distinct(influx_df$sepaSite)
  info$nPens <- 1
  info$IP_penVolume <- farm_i$vol
  params$IP_bg <- params$IP_bg_m3*info$IP_penVolume
  params$IP_halfSat <- params$IP_halfSat_m3*info$IP_penVolume
  params$surv_beta_farms <- map(1:info$nFarms,
                                ~params$surv_beta + 
                                  rbind(rnorm(info$nStages, 
                                              0, 
                                              params$surv_int_farm_sd[1:info$nStages]),
                                        0))
}
```

To run it:

```{r}
#| label: "simulate"
#| warning: false
#| message: false
out_df <- simulate_farm_pops_mn_lpf(params, info, influx_df, 
                                    farm_env, farm_env_daily, out_dir)
saveRDS(farm_env_avg, glue("{out_dir}/farm_env_avg.rds"))
```

In addition to the dataframe returned (daily mean lice per fish: latent (`mu`) and observed(`y`)), all processed inputs, intermediate structures, and output is stored in `{r} out_dir`.

# Results

## Ensemble infestation pressure

```{r}
#| label: "ensIP_fig"
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Ensemble daily copepodid infestation pressure. Farms are ordered from uploch (light) to downloch (dark). Thin lines are individual pens with thick lines as farm-level averages."
#| fig-dpi: 400
#| echo: false
#| message: false
ens_IP <- readRDS(glue("{out_dir}/ensIP.rds"))

if(use_farm_avg) {
  ensIP_df <- expand_grid(sepaSite=farm_i$sepaSite,
                          time=unique(farm_env$time)) |>
    mutate(ensIP=c(ens_IP),
           influx_m3=c(t(apply(ens_IP, 1, function(x) x / info$IP_penVolume))),
           pen=sepaSite) |>
    left_join(farm_i |> select(sepaSite, nPens)) |>
    mutate(influx_m3=influx_m3,
           sepaSite=factor(sepaSite, levels=farm_order))
  ensIP_farm <- ensIP_df
} else {
  ensIP_df <- expand_grid(pen=pen_df$pen, 
                          time=unique(farm_env$time)) |>
    mutate(ensIP=c(ens_IP),
           influx_m3=c(t(apply(ens_IP, 1, function(x) x / info$IP_penVolume)))) |>
    left_join(pen_df |> select(sepaSite, pen), by="pen") |>
    mutate(sepaSite=factor(sepaSite, levels=farm_order))
  ensIP_farm <- ensIP_df |>
    group_by(time, sepaSite) |>
    summarise(ensIP=sum(ensIP)) |>
    ungroup() |>
    left_join(farm_i |> select(sepaSite, nPens)) |>
    mutate(influx_m3=ensIP / ((pi*30^2) * 20 * nPens),
           sepaSite=factor(sepaSite, levels=farm_order))
}

p_ls <- list(
  ggplot(ensIP_df, aes(time, influx_m3, colour=sepaSite)) +
    geom_line(aes(group=pen), linewidth=0.1, alpha=0.5) +
    geom_line(data=ensIP_farm, linewidth=0.5) +
    scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
    scale_y_continuous("Ensemble daily copepodids / m3 / h") +
    scico::scale_colour_scico_d(palette="devon", end=0.8, direction=1, guide="none") +
    facet_grid(.~sepaSite) +
    theme(axis.title.x=element_blank()),
  ggplot(ensIP_df, aes(time, (influx_m3)^0.25, colour=sepaSite)) +
    geom_line(aes(group=pen), linewidth=0.1, alpha=0.5) +
    geom_line(data=ensIP_farm, linewidth=0.5) +
    scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
    scale_y_continuous("(Daily copepodids / m3 / h)^0.25") +
    scico::scale_colour_scico_d(palette="devon", end=0.8, direction=1, guide="none") +
    facet_grid(.~sepaSite) +
    theme(axis.title.x=element_blank()))

cowplot::plot_grid(plotlist=p_ls, ncol=1, align="hv")
```

## Attachment rates

```{r}
#| label: "attach_fig"
#| fig-width: 10
#| fig-height: 3.5
#| fig-cap: "Daily copepodid infestation pressure by simulation. Farms are ordered from uploch (light) to downloch (dark). Thin lines are individual pens with thick lines as farm-level averages."
#| fig-dpi: 400
#| echo: false
pr_attach <- readRDS(glue("{out_dir}/pr_attach.rds"))

if(use_farm_avg) {
  pr_attach_df <- expand_grid(sepaSite=farm_i$sepaSite,
                              time=unique(farm_env$time)) |>
    mutate(pr_attach=c(pr_attach),
           pen=sepaSite) |>
    mutate(sepaSite=factor(sepaSite, levels=farm_order))
  pr_attach_farm <- pr_attach_df
} else {
  pr_attach_df <- expand_grid(pen=pen_df$pen, 
                              time=unique(farm_env$time)) |>
    mutate(pr_attach=c(pr_attach))  |>
    left_join(pen_df |> select(sepaSite, pen), by="pen") |>
    mutate(sepaSite=factor(sepaSite, levels=farm_order))
  pr_attach_farm <- pr_attach_df |>
    group_by(time, sepaSite) |>
    summarise(pr_attach=mean(pr_attach)) |>
    ungroup() |>
    mutate(sepaSite=factor(sepaSite, levels=farm_order))
}


ggplot(pr_attach_df, aes(time, pr_attach, colour=sepaSite)) +
  geom_line(aes(group=pen), linewidth=0.1, alpha=0.5) +
  geom_line(data=pr_attach_farm, linewidth=0.1) +
  scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
  scale_y_continuous("Attachment probability", limits=c(0, 1)) +
  scico::scale_colour_scico_d(palette="devon", end=0.8, direction=1, guide="none") +
  facet_grid(.~sepaSite) +
  theme(axis.title.x=element_blank())
```

## Survival rates

```{r}
#| label: "surv_fig"
#| fig-width: 10
#| fig-height: 7
#| fig-cap: "Daily survival rates by stage. Farms are ordered from uploch (light) to downloch (dark). Thin lines are individual pens with thick lines as farm-level averages."
#| fig-dpi: 400
#| echo: false
pr_surv <- readRDS(glue("{out_dir}/stage_survRate.rds"))

if(use_farm_avg) {
  pr_surv_df <- expand_grid(stage=c("chal", "prea", "ad"),
                            date=unique(farm_env_daily$date),
                            sepaSite=farm_i$sepaSite) |>
    mutate(pr_surv=c(pr_surv),
           pen=sepaSite,
           sepaSite=factor(sepaSite, levels=farm_order),
           stage=factor(stage, levels=c("chal", "prea", "ad"))) 
  pr_surv_farm <- pr_surv_df
} else {
  pr_surv_df <- expand_grid(stage=c("chal", "prea", "ad"),
                            date=unique(farm_env_daily$date),
                            pen=pen_df$pen) |>
    mutate(pr_surv=c(pr_surv))  |>
    left_join(pen_df |> select(sepaSite, pen), by="pen") |>
    mutate(sepaSite=factor(sepaSite, levels=farm_order),
           stage=factor(stage, levels=c("chal", "prea", "ad")))
  pr_surv_farm <- pr_surv_df |>
    group_by(date, sepaSite) |>
    summarise(pr_surv=mean(pr_surv)) |>
    ungroup() |>
    mutate(sepaSite=factor(sepaSite, levels=farm_order))
}


ggplot(pr_surv_df, aes(date, pr_surv, colour=sepaSite)) +
  geom_line(aes(group=pen), linewidth=0.1, alpha=0.5) +
  geom_line(data=pr_surv_farm, linewidth=0.1) +
  scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
  scale_y_continuous("Survival probability", limits=c(0, 1)) +
  scico::scale_colour_scico_d(palette="devon", end=0.8, direction=1, guide="none") +
  facet_grid(stage~sepaSite) +
  theme(axis.title.x=element_blank())
```

## Lice per fish

We can look at both the latent mean lice per fish (`mu`) and the observed mean lice per fish (`y`). The latent mean is recorded each day, with observations taken once each week on the corresponding day (`farm_i$firstDay`) as the average of `farm_env$nFishSample[farm,day]` fish.

```{r}
#| label: "lice_fig1"
#| fig-width: 10
#| fig-height: 15
#| fig-cap: "Daily mean latent lice per fish (mu)."
#| fig-dpi: 400
#| echo: false
#| message: false

mu_df <- out_df |>
  select(date, sepaSite, pen, treat, starts_with("mu_")) |>
  pivot_longer(c(starts_with("mu_"))) |>
  mutate(type=str_split_fixed(name, "_", 2)[,1],
         stage=str_split_fixed(name, "_", 2)[,2]) |>
  mutate(stage=factor(stage, levels=c("chal", "prea", "am", "af"))) 
ybar_df <- out_df |>
  filter(sampled) |>
  # mutate(across(starts_with("ybar_"), ~.x/nFishSampled)) |>
  select(date, sepaSite, pen, treat, starts_with("ybar_")) |>
  pivot_longer(c(starts_with("ybar_"))) |>
  mutate(type=str_split_fixed(name, "_", 2)[,1],
         stage=str_split_fixed(name, "_", 2)[,2]) |>
  mutate(stage=factor(stage, levels=c("chal", "prea", "am", "af"))) 
y_df <- out_df |>
  filter(sampled) |>
  # mutate(across(starts_with("y_"), ~.x/nFishSampled)) |>
  select(date, sepaSite, pen, treat, starts_with("y_")) |>
  pivot_longer(c(starts_with("y_"))) |>
  mutate(type=str_split_fixed(name, "_", 2)[,1],
         stage=str_split_fixed(name, "_", 2)[,2]) |>
  mutate(stage=factor(stage, levels=c("chal", "prea", "am", "af"))) 

ggplot(mu_df, aes(date)) + 
  geom_line(aes(y=value, colour=stage, group=paste(stage, pen))) +
  geom_rug(data=out_df |> filter(treat==1), sides="b", colour="black") +
  scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
  scale_colour_brewer(type="qual", palette="Paired") +
  labs(y="Latent mean lice per fish") +
  facet_grid(sepaSite~., scales="free_y") + 
  theme(axis.title.x=element_blank())
```

```{r}
#| label: "lice_fig2"
#| fig-width: 10
#| fig-height: 15
#| fig-cap: "Population composition."
#| fig-dpi: 400
#| echo: false
#| message: false
mu_df |>
  ggplot(aes(date)) +
  geom_area(aes(y=value, fill=stage), position="fill", colour="grey30", linewidth=0.1) +
  geom_rug(data=out_df |> filter(treat==1), sides="b", colour="black") +
  scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
  scale_fill_brewer(type="qual", palette="Paired") +
  labs(y="Proportion of population") +
  facet_grid(sepaSite~., scales="free_y") + 
  theme(axis.title.x=element_blank(),
        legend.position="bottom")
if(save_figs) {
  ggsave("admin/project_meetings/figs_temp/dataExamp_lice_popDistr.png", width=5, height=7)
}
```

```{r}
#| label: "lice_fig3"
#| fig-width: 10
#| fig-height: 10
#| fig-cap: "Daily mean lice per fish, including latent (mu, lines) and observed (y, points)."
#| fig-dpi: 400
#| echo: false
ggplot(mu_df, aes(date)) + 
  geom_line(aes(y=value, group=pen), colour="cadetblue", alpha=0.5) +
  geom_jitter(data=y_df, aes(y=value), shape=1, size=0.8, width=0.2, height=0) +
  geom_rug(data=out_df |> filter(treat==1), sides="b", colour="black") +
  scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
  labs(y="Lice per fish") +
  facet_grid(stage~sepaSite, scales="free_y") + 
  theme(axis.title.x=element_blank())
if(save_figs) {
  ggsave("admin/project_meetings/figs_temp/dataExamp_lice_mu_y.png", width=7, height=6)
}
```

```{r}
#| label: "lice_fig4"
#| fig-width: 10
#| fig-height: 10
#| fig-cap: "Daily mean lice per fish for adult female stages (af: all, gr: gravid, ng: non-gravid), including latent (mu, lines) and observed (y, points)."
#| fig-dpi: 400
#| echo: false

mu_farm <- mu_df |>
  group_by(date, sepaSite, type, stage) |>
  summarise(value=mean(value)) |>
  ungroup()
y_farm <- y_df |>
  group_by(date, sepaSite, type, stage) |>
  summarise(value=mean(value)) |>
  ungroup()
af <- c("af")
ggplot(mu_df |> filter(stage %in% af), aes(date)) + 
  geom_hline(yintercept=0.5, linetype=2) + 
  geom_line(aes(y=value, group=pen), colour="cadetblue", alpha=0.5, linewidth=0.1) +
  geom_line(data=mu_farm |> filter(stage %in% af), aes(y=value),
            colour="cadetblue", linewidth=0.9) +
  geom_jitter(data=y_df |> filter(stage %in% af), aes(y=value),
              shape=1, size=0.8, width=0.2, height=0) +
  geom_point(data=y_farm |> filter(stage %in% af), aes(y=value), shape=4, size=2) +
  geom_rug(data=out_df |> filter(treat==1), sides="b", colour="black") +
  scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
  labs(y="Lice per fish") +
  facet_grid(sepaSite~stage, scales="free_y") + 
  theme(axis.title.x=element_blank())
```

```{r}
#| label: "lice_fig5"
#| fig-width: 15
#| fig-height: 5
#| fig-cap: "Daily mean lice per fish for adult female stages (af: all, gr: gravid, ng: non-gravid), including latent (mu, lines) and observed (y, points)."
#| fig-dpi: 400
#| echo: false
ggplot(mu_df |> filter(stage %in% af), aes(date)) + 
  geom_hline(yintercept=0.5, linetype=2) + 
  geom_line(aes(y=value, group=pen), colour="cadetblue", alpha=0.5, linewidth=0.1) +
  geom_line(data=mu_farm |> filter(stage %in% af), aes(y=value),
            colour="cadetblue", linewidth=0.9) +
  # geom_jitter(data=y_df |> filter(stage %in% af), 
  #             shape=1, size=0.8, width=0.2, height=0) +
  geom_point(data=y_farm |> filter(stage %in% af), aes(y=value), shape=4, size=2) +
  geom_rug(data=out_df |> filter(treat==1), sides="b", colour="black") +
  scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
  labs(y="Simulated mean AF lice per fish\n(line: population;  points: counts)") +
  facet_grid(.~sepaSite) + 
  theme(axis.title.x=element_blank())
if(save_figs) {
  ggsave("admin/project_meetings/figs_temp/dataExamp_lice_mu_y_AF.png", width=9, height=4)
}
```

```{r}
#| label: "lice_fig6"
#| fig-width: 10
#| fig-height: 5
#| fig-cap: "Daily mean lice per fish for adult female stages (af: all, gr: gravid, ng: non-gravid), including latent (mu, lines) and observed (y, points)."
#| fig-dpi: 400
#| echo: false
y_df |> 
  filter(stage != "am") |> 
  ggplot(aes(date)) + 
  geom_hline(yintercept=0.5, linetype=2) + 
  geom_point(aes(y=value), shape=4) + 
  geom_rug(data=out_df |> filter(treat==1), sides="b", colour="black") +
  scale_x_date(date_breaks="6 months", date_labels="%b", date_minor_breaks="2 month") +
  labs(y="Observed lice per fish") +
  facet_grid(stage~sepaSite, scales="free_y") + 
  theme(axis.title.x=element_blank())
if(save_figs) {
  ggsave("admin/project_meetings/figs_temp/dataExamp_lice_y.png", width=10, height=7)
}
```

Note that these are simulated with arbitrary and randomized model parameters and would not be expected to fully conform with observed values.

# References
