---
title: "E(tive)Lice verification assessment"
author: "Tim Szewczyk"
format: html
theme: litera
mainfont: Garamond
monofont: Lucida Console
editor: visual
toc: true
number-sections: true
bibliography: references.bib
editor_options: 
  chunk_output_type: console
---

# Overview

This document evaluates the IEM performance on the simulated verification data sets.

# Setup

```{r}
#| message: false
#| warning: false
#| results: false
#| echo: false
library(tidyverse)
library(sf)
library(glue)
library(here)
proj_dir <- here()
theme_set(theme_bw())
dir(glue("{proj_dir}/code/fn"), ".R", full.names=T) |> walk(source)

farm_order <- c("FFMC84", "FFMC32", "APT1", "SAR1")#, "FFMC27")
sim_dir <- glue("{proj_dir}/data/sim/")
```

```{r}
#| echo: false
stages <- c("Ch", "PA", "Ad", "Gr")
stage_trans <- c("Ch-PA", "PA-Ad", "Ad-Gr")
param_key <- tibble(name=c(paste0("attach_beta[", 1:5, "]"),
                           paste0("ensWts_p[", 1:6, "]"),
                           "IP_bg", "IP_bg_m3",
                           paste0("surv_beta[1,", 1:4, "]"),
                           paste0("surv_beta[2,", 1:4, "]"),
                           paste0("pMoltF_beta[1,", 1:3, "]"),
                           paste0("pMoltF_beta[2,", 1:3, "]"),
                           paste0("pMoltM_beta[1,", 1:2, "]"),
                           paste0("pMoltM_beta[2,", 1:2, "]"),
                           paste0("thresh_GDD[", 1:3, ",1]"),
                           paste0("thresh_GDD[", 1:2, ",2]"),
                           "lifespan",
                           paste0("detect_p[", 1:4, "]"),
                           "nb_prec"),
                    label=c(paste0("attach_", c("Int", "RW", "Sal", "UV", "UVsq")),
                            paste0("ensWt_", 1:6),
                            "IP_bg", "IP_bg_m3",
                            paste0("surv_Int_", stages),
                            paste0("surv_Sal_", stages),
                            paste0("pMoltF_Int_", stage_trans),
                            paste0("pMoltF_Temp_", stage_trans),
                            paste0("pMoltM_Int_", stage_trans[-3]),
                            paste0("pMoltM_Temp_", stage_trans[-3]),
                            paste0("moltF_GDD_", stage_trans),
                            paste0("moltM_GDD_", stage_trans[-3]),
                            "lifespan_GDD",
                            paste0("p_detect_", stages),
                            "neg_binom_prec"
                    )) |>
  mutate(label=factor(label, levels=unique(label)))
```

# Load data and posterior summaries

```{r}
#| message: false
#| warning: false
out_parts_df <- dir(sim_dir, "sim_")[1:6] |>
  map_dfr(
    ~full_join(
      bind_rows(read_csv(glue("{sim_dir}/{.x}/posterior_summary_pt1.csv"), 
                         show_col_types=F) |> 
                  filter(name != "nb_prec"),
                read_csv(glue("{sim_dir}/{.x}/posterior_summary_pt2_pop.csv"), 
                         show_col_types=F) |>
                  filter(name != "nb_prec"),
                read_csv(glue("{sim_dir}/{.x}/posterior_summary_pt3.csv"), 
                         show_col_types=F)),
      read_csv(glue("{sim_dir}/{.x}/params_full.csv"), show_col_types=F), 
      by=join_by(name)
    ) |>
      mutate(sim=paste0("dat_", str_sub(.x, -2, -1))) |>
      inner_join(param_key, by=join_by(name)))
out_full_df <- dir(sim_dir, "sim_")[1:6] |>
  map_dfr(
    ~full_join(
      read_csv(glue("{sim_dir}/{.x}/posterior_summary_full_FEMALE.csv"), show_col_types=F),
      read_csv(glue("{sim_dir}/{.x}/params_full.csv"), show_col_types=F), 
      by=join_by(name)
    ) |>
      mutate(sim=paste0("dat_", str_sub(.x, -2, -1))) |>
      inner_join(param_key, by=join_by(name)))
out_df <- bind_rows(
  out_parts_df |> mutate(fit="by_part"),
  out_full_df |> mutate(fit="in_full")
)
```

## IP and attachment

### Ensemble IP weights

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 3
#| echo: false
p_ensWts <- out_df |>
  filter(grepl("ensWts", name)) |>
  verification_plot(scales="free_x") +
  scale_x_continuous(limits=c(0, 1), breaks=round((0:3)/3, 2))
p_ensWts
```

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 2
#| echo: false
out_df |>
  filter(grepl("ensWts", name)) |>
  mutate(inCI=between(value, L025, L975)) |>
  pr_in_CI_plot()
```

```{r}
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 3
#| echo: false
out_df |>
  filter(grepl("ensWts", name)) |>
  mutate(inCI=between(value, L025, L975)) |>
  scatter_post_mean_plot()
```

### Background IP

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 3
#| echo: false
p_IP <- out_df |> 
  filter(grepl("IP_bg", name)) |>
  verification_plot(ncol=3, scales="free")
p_IP
```

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 2
#| echo: false
out_df |>
  filter(grepl("IP_bg", name)) |>
  mutate(inCI=between(value, L025, L975)) |>
  pr_in_CI_plot()
```

```{r}
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 3
#| echo: false
out_df |>
  filter(grepl("IP_bg_m3", name)) |>
  mutate(inCI=between(value, L025, L975)) |>
  scatter_post_mean_plot()
```

### P(attach)

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 3
#| echo: false
p_attach <- out_df |>   
  filter(grepl("attach_beta", name)) |>   
  verification_plot(scales="free_x") +     
  geom_vline(xintercept=0, linetype=3) 
p_attach
```

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 2
#| echo: false
out_df |>
  filter(grepl("attach_beta", name)) |>
  mutate(inCI=between(value, L025, L975)) |>
  pr_in_CI_plot()
```

```{r}
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 3
#| echo: false
out_df |>
  filter(grepl("attach_beta", name)) |>
  mutate(inCI=between(value, L025, L975)) |>
  scatter_post_mean_plot()
```

## Development and survival

### P(survival)

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 6
#| echo: false
p_surv <- out_df |> filter(grepl("surv_beta", name)) |>
  verification_plot(ncol=4, scales="free_x") +
  geom_vline(xintercept=0, linetype=3)
p_surv
```

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 2
#| echo: false
out_df |>
  filter(grepl("surv_beta", name)) |>
  mutate(inCI=between(value, L025, L975)) |>
  pr_in_CI_plot()
```

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 2
#| echo: false
out_df |>
  filter(grepl("surv_beta", name)) |>
  mutate(inCI=between(value, L025, L975)) |>
  scatter_post_mean_plot()
```

### P(molt) intercept

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 3
#| echo: false
p_pMoltInt <- out_df |> 
  filter(grepl("pMolt.*_beta\\[1", name)) |>
  mutate(across(where(is.numeric), plogis)) |>
  verification_plot(ncol=5) +
  geom_vline(xintercept=0, linetype=3) +
  scale_x_continuous(limits=c(0, 0.1), breaks=c(0, 0.05, 0.1))
p_pMoltInt
```

### P(molt) temperature effect

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 3
#| echo: false
p_pMoltTemp <- out_df |> 
  filter(grepl("pMolt.*_beta\\[2", name)) |>
  verification_plot(ncol=5) +
  geom_vline(xintercept=0, linetype=3)
p_pMoltTemp
```

## Sampling and detection

### P(detect)

```{r}
#| fig-width: 8
#| fig-height: 3
#| echo: false
p_detectp <- out_df |> 
  filter(grepl("detect_p", name)) |>
  verification_plot(scales="free") +
  xlim(0, 1)
p_detectp
```

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 2
#| echo: false
out_df |>
  filter(grepl("detect_p", name)) |>
  mutate(inCI=between(value, L025, L975)) |>
  pr_in_CI_plot()
```

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 2
#| echo: false
out_df |>
  filter(grepl("detect_p", name)) |>
  mutate(inCI=between(value, L025, L975)) |>
  scatter_post_mean_plot()
```

### Negative binomial precision

```{r}
#| fig-width: 4
#| fig-height: 3
#| echo: false
p_nb <- out_df |> 
  filter(grepl("nb_prec", name)) |>
  verification_plot()
p_nb
```

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 2
#| echo: false
out_df |>
  filter(grepl("nb_prec", name)) |>
  mutate(inCI=between(value, L025, L975)) |>
  pr_in_CI_plot()
```

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 2
#| echo: false
out_df |>
  filter(grepl("nb_prec", name)) |>
  mutate(inCI=between(value, L025, L975)) |>
  scatter_post_mean_plot()
```

## All parameters

```{r}
full_sum_df <- out_df |>
  filter(fit=="in_full") |>
  filter(name != "IP_bg") |>
  mutate(name=if_else(name=="lifespan", "thresh_GDD[,4]", name)) |>
  mutate(param_type=str_split_fixed(name, "\\[", 2)[,1]) 
lims_df <- full_sum_df |>
  group_by(param_type) |>
  summarise(lower=min(L25, value, na.rm=T),
            upper=max(L75, value, na.rm=T))
full_sum_df |>
  ggplot(aes(mn, value)) +
  geom_point(data=lims_df, aes(lower, lower), colour="white", size=0.2) +
  geom_point(data=lims_df, aes(upper, upper), colour="white", size=0.2) +
  geom_abline(linetype=3) +
  geom_point(shape=1, alpha=0.75) +
  geom_linerange(aes(xmin=L10, xmax=L90)) +
  facet_wrap(~param_type, scales="free", nrow=2) +
  # coord_(xlim=lims, ylim=lims) +
  labs(x="Posterior mean + 80% CI", 
       y="True value",
       title="Verification across 6 simulated datasets")
ggsave("admin/project_meetings/figs_temp/full_performance.png", width=8, height=4.5, dpi=300)
```

# Posterior regressions

## Survival

```{r}
S_range <- range(readRDS("data/sim/sim_01/sal_mx.rds"))
S_df <- tibble(sal=seq_range(S_range, length.out=100))
pSurv_df <- out_df |>
  filter(sim %in% out_full_df$sim) |>
  filter(grepl("surv_beta", name)) |>
  mutate(stage=str_sub(label, -2, -1),
         beta=paste0("b", str_sub(name, -4, -4))) |>
  select(stage, beta, mn, value, sim, fit) |>
  pivot_longer(3:4, names_to="src", values_to="value") |>
  filter(! (src=="value" & fit=="by_part")) |>
  mutate(fit=if_else(src=="value", "True", fit)) |>
  select(-src) |>
  pivot_wider(names_from="beta", values_from="value") |>
  mutate(S_df=list(S_df)) |>
  unnest(S_df) |>
  mutate(pSurv=plogis(b1 + b2*sal),
         salinity=sal+30,
         stage=factor(stage, levels=unique(stage)))
```

```{r}
pSurv_df |>
  ggplot(aes(salinity, pSurv, group=paste(stage, fit), colour=stage, linetype=fit)) + 
  geom_line() +
  scale_colour_manual(values=RColorBrewer::brewer.pal(n=6, name="Paired")[c(1,2,4,6)]) +
  scale_linetype_manual(values=c(3, 2, 1)) +
  ylim(0, 1) + 
  labs(x="Salinity (psu)", y="Daily Pr(Survival)") +
  # facet_wrap(~sim)
  facet_grid(stage~sim)
```

```{r}
pSurv_df |>
  ggplot(aes(salinity, pSurv, group=paste(stage, fit), colour=stage, linetype=fit)) + 
  geom_line() +
  # scale_colour_manual("Fit", values=c("#377eb8", "#4daf4a")) +
  ylim(0, 1) + 
  labs(x="Salinity (psu)", y="Daily Pr(Survival)") +
  facet_wrap(~sim)
```

## Development rate

```{r}
T_range <- range(readRDS("data/sim/sim_01/temp_mx.rds"))
T_z_range <- range(readRDS("data/sim/sim_01/temp_z_mx.rds"))

T_df <- tibble(temp=seq_range(T_range, length.out=100),
               temp_z=seq_range(T_z_range, length.out=100))
pMolt_df <- out_df |>
  filter(sim %in% out_full_df$sim) |>
  filter(grepl("pMolt.*_beta\\[", name)) |>
  mutate(sex=if_else(grepl("F", name), "F", "M"),
         stage=str_sub(label, -5, -1),
         beta=paste0("b", str_sub(name, -4, -4))) |>
  select(sex, stage, beta, mn, sim, fit) |>
  pivot_wider(names_from="beta", values_from="mn") |>
  mutate(T_df=list(T_df)) |>
  unnest(T_df) |>
  mutate(pMolt=plogis(b1 + b2*temp_z),
         mnDays=1/pMolt,
         stage=factor(stage, levels=unique(stage))) 
```

```{r}
pMolt_df |>
  ggplot(aes(temp, pMolt, group=paste(sim, fit), colour=fit)) + 
  geom_line(alpha=0.3) +
  scale_colour_manual("Fit", values=c("#377eb8", "#4daf4a")) +
  ylim(0, 1) + 
  labs(x="Temperature (C)", y="Daily Pr(Molt)") +
  facet_grid(sex~stage)
```

```{r}
pMolt_df |>
  ggplot(aes(temp, mnDays, group=paste(sim, fit), colour=fit)) + 
  geom_line(alpha=0.3) +
  scale_colour_manual("Fit", values=c("#377eb8", "#4daf4a")) +
  ylim(0, NA) +
  labs(x="Temperature (C)", y="Mean days per stage") +
  facet_grid(sex~stage)
```
