---
title: "Kragesteen 2023"
author: "Tim Szewczyk"
date: "2024-12-02"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse); theme_set(theme_bw())
```

# Population model outline

## Overview

This is an overview of the model from [Kragesteen et al. 2013](https://doi.org/10.1016/j.aquaculture.2023.739759). This was focused on production zone 2 in Norway (narrow band south of Bergen) with 49 farms from 2016-2019. Seawater temperature at 3m varies from 4-6C in winter to 15-18C in the summer. The innermost part of the fjord arms have salinity below 20 ppt, with higher salinity below 5m and beyond the most protected areas. The biophysical model was barely described, but should be the Norwegian model given the location and co-authors.

## Larvae production

$l_i(t) = {AG}_i(t) \cdot E(T_i(t)) \cdot \Delta t$

with $l$ larvae produced at site $i$ and time $t$, gravid adult females $AG$, egg production $E$ at temperature $T$, and $\Delta t$ set to 0.5 days. The effect of temperature on egg production was a linear interpolation (*NOT* a linear model! This is strange, right?), using Stien (2005) and Samsing (2016) for 3-5C and Hamre (2019) for 6-18C. Production at 20C was set to 90.8 eggs per female per day, identical to 18C (so flatline after 18C?).

```{r}
egg_df <- tibble(eggs_AG_day=c(8.4, 26.6, 28.6, 80.9, 90.8, 90.8),
                 temperature=c(3, 5, 6, 12, 18, 20))
ggplot(egg_df, aes(temperature, eggs_AG_day)) + 
  geom_point() + geom_line() +
  geom_line(stat="smooth", method="lm", formula=y~x+I(x^2), 
            linewidth=0.5, colour="steelblue") +
  scale_x_continuous(breaks=seq(0, 20, by=2)) +
  scale_y_continuous(breaks=seq(0, 100, by=10)) 
```

This is quite a bit higher than the 28.2 eggs per day assumed by biotracker defaults. From Moriarty (2024):

> The production of eggs per ovigerous female louse is more variable compared with egg or embryo viability (Johnson & Albright 1991, Brooker et al. 2018). **For modelling applications, rates of between 28 and 30 viable eggs per day are used**, as a multiplication factor (for each adult female) and are applied as a default (e.g. Adams et al. 2021, Murray & Moriarty 2021) or may be specified as a function of temperature (e.g. Sandvik et al. 2020). However, this is an area where more detailed empirical data, following the approach by á Norði et al. (2016) or similar, and subsequent modelling could be useful, particularly if freshwater intrusions from nearby rivers occur in the modelled area.

This discrepancy seems pretty large since the 28-30 range occurs around 6C which is on the low end of what occurs in Scotland. To code this into biotracker for testing, I'd need to fit a function. A quadratic seems pretty reasonable. I'd need to confirm with the temperature ranges within WeStCOMS, though in current climate, it's unlikely that the water ever gets much warmer or colder.

## Attachment

Attachment is a function of the copepodid infection pressure and the receiving area ratio:

$A_{i,q}(t_0) = R(W_i) \cdot \Big( L_{0,i}(t) + \sum_j \sum_d C_{i,j,d}(t) \cdot l_j (t-d) \Big)$

where $A$ is the abundance of cohort $q$, $R$ is the receiving area ratio, $L$ is the background IP from wild fish, and $C$ is the copepodid infection pressure from farm $j$. The copepodid infection pressure incorporates a background rate $L_{0,i}(t)$ which somehow is a function of time and location. In at least some of the simulations, $L = 30,000 ~ \text{lice d}^{-1}$ for the first 200 days, then set to 0. The infection pressure from other farms is normal, with daily connectivity between farms.

The receiving area ratio is somewhat stranger. This is the overall proportion of copepodids that are expected to attach, which can be interpreted as a per-louse attachment rate. There is consequently an expectation that it depends on the number and size of fish. They evaluated three different methods. First, a constant ratio of lice attach to a farm independent of everything:

$R(W_i) = r$

Second, a type II functional response where lice attachment was dependent on biomass and max biomass, with a scaling parameter $a = 100$ and maximum surface ratio $r = 0.15$:

$R(W_i) = r \frac{\frac{W}{W_max} a}{1 + \frac{W}{W_max} a}$

Third, attachment proportion is based on the biomass $W$, stocking density $D$, and receiving area $V_r$:

$R(W_i) = \sqrt{\frac{W}{DV_r}}$

See Fig. 3 for idealized curves, and Fig. 7 for results.

None of these depend on anything but the fish. This is a piece that I think could be adjusted.

They calculate R^2^ for each method (predicted vs observed abundance, AF/PAAM/AF+PAAM; PAAM=pre-adult + adult male; AF=adult female incl. gravid) and report that method three was the best. It generally was better (increase in R^2^ of 0.08 to 0.16) though not for AF in period 2. There was basically no difference between 1 and 2.

## Biological age and mortality

The change in biological age for each cohort depended on temperature following Hamre (2019) such that:

$B(t + \Delta t) = B(t) + \frac{b T_i^2 + c T_i + d}{5 H_s} \Delta t$

where $b$, $c$, and $d$ are fitted constants, and $H_s$ is the stage fraction. **What is the stage fraction?** Read Hamre (2019) in more detail. The quadratic parameters were separate for males and females. The structure of these equations

Mortality was implemented as:

$A_{i,q}(t + \Delta t) = A_{i,q}(t) \cdot e^{-\mu_s \Delta t} \cdot \psi_{i,s}(t)$

with stage and sex specific mortality rate $\mu$ and treatment mortality $\psi$. In practice, the mortality rates only differed by stage.
